Ab Initio Effective One-Electron Potentials Revisited: General Theory and Example Applications for Charge-Transfer Energy Calculations

### 2019-05-24 papers/2_oep_technique/paper-r0.tex
Working formulae derived for DF-temr in CT energy of Murrell et al

\section{\label{s:2}Effective One-Electron Potentials: Configuration Interaction}

The time\hyp{}independent Schr{\"o}dinger equation provides the stationary solution to the
electronic state
of a molecular system in terms of its $N$\hyp{}electron wavefunction, $\Psi$,
%
\begin{equation} \label{e:schrodinger}
 E = \tBraKet{\Psi}{\mathscr{H}}{\Psi} \;,
\end{equation}
%
for a quantum Hamiltonian, $\mathscr{H} = \hat{h}(1) + \hat{o}(2)$,
with $\hat{h}(1)$ being the core Hamiltonian one\hyp{}electron operator
and $\hat{o}(2)$ the two\hyp{}electron repulsion operator.
From the above solution, any property $P$, understood as an expectation value of the
associated operator $\mathscr{P}$, can be computed according to
%
\begin{equation} \label{e:prop}
 P = \tBraKet{\Psi}{\mathscr{P}}{\Psi} \;.
\end{equation}
%
The exact solution to Eq.~\eqref{e:schrodinger} is formally given by
the configuration interaction (CI) expansion around the reference 
(approximate) wavefunction, $\Psi^{(0)}$,
typically the solution to the Hartree\hyp{}Fock (HF) problem,
%
\begin{equation} \label{e:ci}
 \Ket{\Psi} = \Ket{\Psi^{(0)}} + \sum_{ra} C_{r}^{a} \Ket{\Psi_{r}^{a}} + 
	 \sum_{rsab} C_{rs}^{ab} \Ket{\Psi_{rs}^{ab}} + \ldots
\end{equation}
%
In the above equation,
$\Ket{\Psi_{rs\ldots}^{ab\ldots}}$ are called the CI configurations
with the associated CI coefficients $C_{rs\ldots}^{ab\ldots}$, created by `exciting'
one or more electrons from the $a$th ($b$th) occupied molecular orbital to the $a$th ($b$th)
virtual (unoccupied) molecular orbital, obtained from the HF solution.
By inserting Eq.~\eqref{e:ci} into Eq.~\eqref{e:prop} 
and using the Slater\hyp{}Condon rules for evaluating the one\hyp{} and two\hyp{}electron
operator matrix element between
the CI configurations, it can be shown that the following holds
%
\begin{equation} \label{e:exp-val-series}
	P =
	\sum_{ij} P_{ij} \tBraKet{i}{\hat{h}}{j}
	+ \sum_{ijkl} P_{ijkl} \BraKet{ij}{kl}  \;,
\end{equation}
%
%where $\tBraKet{i}{\hat{h}}{j} \equiv \int \phi^*_i({\bf r}) \hat{h} \phi_j({\bf r}) d{\bf r} $
%$\BraKet{ij}{kl}$ is the electron\hyp{}repulsion integral (ERI) defined by
%%
%\begin{equation}
%	\BraKet{ij}{kl} \equiv
%	\iint 
%	\frac{ \phi_i^{*}({\bf r}_1) \phi_j({\bf r}_1) 
%	       \phi_k^{*}({\bf r}_2) \phi_l({\bf r}_2) }{ \vert {\bf r}_1 - {\bf r}_2 \vert}
%	d{\bf r}_1 d{\bf r}_2  \;,
%\end{equation}
%%
where the second\hyp{}rank tensor ${\bf P}^{(2)}$ 
and the fourth\hyp{}rank tensor ${\bf P}^{(4)}$ 
are well defined and characteristic for the property of interest.
In fact, they are certain functions of the CI coefficients, i.e.,
%
\begin{equation}
{\bf P} = {\bf P}
\left(
 \left\{
  C_{r}^{a}, C_{rs}^{ab}, \ldots 
 \right\} 
\right)
\end{equation}
%
and, in principle, are known as long as the solution to the CI problem 
or its approximation is available.

@@

\subsection{\label{ss:2.3}Defining Effective Potentials}

@@

\subsection{\label{ss:2.4}Interaction-Induced Property: Effective Potentials}

@@

\subsection{\label{s.334}Density Fitting of OEP's}

@@

An arbitrary one-electron potential of molecule $A$ acting on any state vector 
associated with molecule $A$ can be expanded in an auxiliary space centered 
on $A$ as
\begin{equation}
   \hat{v}^{A}\Ket{i} = \sum_{\xi\eta} \hat{v}^{A}\Ket{\xi} [{\bf S}^{-1}]_{\xi\eta} \BraKet{\eta}{i}
\end{equation}
under the necessary assumption that the auxiliary basis set is nearly complete,
i.e., 
$\sum_{\xi\eta} \Ket{\xi}[{\bf S}^{-1}]_{\xi\eta} \Bra{\eta} \cong 1$.
%In a special case when the basis set is orthogonal (e.g., molecular orbitals)
%the above relation simplifies to
%\begin{equation}
%   v\vert i) = \sum_{\xi} v\vert \xi) ( \xi \vert i)
%\end{equation}
The above general and exact expansion can be obtained by 
utilizing the resolution of identity or, equivalently, density fitting
in the complete space,
%
\begin{equation}
 \hat{v}^{A}\Ket{i} = \sum_{\xi} V^A_{i\xi} \Ket{\xi} \;.
\end{equation}
%
In the above equation,
the matrix ${\bf V}^A$
is the representation of the OEP operator $\hat{v}^{A}$
in the complete basis $\{ \varphi_\xi \}$.
By constructing the least\hyp{}squares objective function 
%
\begin{equation}
 Z[{\bf V}^A] = \int 
                     \left[    v({\bf r}_1) \phi_i({\bf r}_1) - \sum_\xi V^A_{i\xi} \varphi_\xi({\bf r}_1) \right]^2
d{\bf r}_1
\end{equation}
%
and requiring that
%
\begin{equation}
 \frac{\partial Z[{\bf V}^A]}{\partial V^A_{i\mu}} = 0 \text{ for all $\mu$}
\end{equation}
%
one finds the coefficients of the $i$th row of ${\bf V}^A$ to be
%
\begin{equation}
% {\bf v}_i = {\bf a}^{(i)} \cdot {\bf S}^{-1}
  V^A_{i\xi} = \sum_\eta [{\bf S}^{-1}]_{\xi\eta} a^{(i)}_\eta \;,
\end{equation}
%
where
%
\begin{subequations}
\begin{align}
 a^{(i)}_\eta &= \int \varphi^*_\eta({\bf r}) \hat{v}^A \phi_i({\bf r}) d{\bf r} \;,\\  
 S_{\eta\xi}  &= \int \varphi^*_\eta({\bf r}) \varphi_\xi({\bf r}) d{\bf r} \;.
\end{align}
\end{subequations}
%
%Since matrix elements of an OEP operator in auxiliary space can be computed 
%in the same way as the matrix elements with any other basis function, 
%one can formally write the following identity
%\begin{equation}
% (X \vert v\vert i) = \sum_{\xi\eta} S_{X \xi} [{\bf S}^{-1}]_{\xi\eta} (\eta\vert v\vert i)
%\end{equation}
%where $ X $ is an arbitrary orbital.
%When the other orbital
%does not belong to molecule $A$ but to the (changing) environment, it is 
%straightforward to compute the resulting matrix element, which is simply given as  
%\begin{equation}
%   (j_{\in B} \vert v^A \vert i_{\in A}) = \sum_\xi {S_{j\xi}} {G_{i\xi}}
%\end{equation}
%where $j$ denotes the other (environmental) basis function.
%
The working formula for $a^{(i)}_\eta$ can be found by applying 
potential for from Eq.~\eqref{e:v-eff}
along with the spectral representation of the effective density from Eq.~\eqref{e:d-spectral} 
which gives
%
\begin{equation}
 a^{(i)}_\eta = \sum_{x\in A} W_{\eta i}^{(x)} + 
 \sum_{\alpha\beta} D^A_{\alpha\beta} 
  \BraKet{\alpha\beta}{\eta i} \;.
%  \iint \frac{ \phi_\alpha^*({\bf r}_1)  \phi_\beta({\bf r}_1) \varphi^*_\eta({\bf r}_2) \phi_i({\bf r}_2)}
% {\vert {\bf r}_1 - {\bf r}_2 \vert }
% d{\bf r}_1 d{\bf r}_2
\end{equation}
%

@@

Density fitting scheme from previous section has practical disadvantage of a nearly-complete basis set
being usually very large (spanned by large amount of basis set vectors). Any non-complete basis set
won't work in the previous example. Since most of basis sets used in quantum chemistry do not form a nealy complete
set, it is beneficial to design a modified scheme in which it is possible to obtain the effective 
matrix elements of the OEP operator in a incomplete auxiliary space. This can be achieved by minimizing 
the following objective function
\begin{multline}
	Z[{\bf V}^A] = \iint 
        \Bigg\{
                     \left[    v^A({\bf r}_1) \phi_i({\bf r}_1) - \sum_\xi V^A_{i\xi} \varphi_\xi({\bf r}_1) \right]
                  \\ \times \frac{1}{\vert {\bf r}_1 - {\bf r}_2 \vert}  
                     \left[    v^A({\bf r}_2) \phi_i({\bf r}_2) - \sum_\xi V^A_{i\xi} \varphi_\eta({\bf r}_1) \right]
		    \Bigg\}
         d{\bf r}_1 d{\bf r}_2  \;.
\end{multline}
Requesting that 
\begin{equation}
 \frac{\partial Z[{\bf V}^A]}{\partial V^A_{i\mu}} = 0 \text{ for all $\mu$}
\end{equation}
one obtains
\begin{equation}
% {\bf G}^{(i)} = {\bf b}^{(i)} \cdot {\bf A}^{-1}
  V^A_{i\xi} = \sum_\eta [{\bf A}^{-1}]_{\xi\eta} b^{(i)}_\eta \;,
\end{equation}
where 
\begin{subequations}
\begin{align}
 b^{(i)}_\eta &= \iint 
                       \frac{ \varphi^*_\eta({\bf r}_1) \hat{v} \phi_i({\bf r}_2) } 
                            {\vert {\bf r}_1 - {\bf r}_2\vert}  
                 d{\bf r}_1 d{\bf r}_2 \;, \\
 A_{\eta\xi}  &= \iint 
                       \frac{ \varphi^*_\eta({\bf r}_1) \varphi_\xi({\bf r}_2) } 
                            {\vert {\bf r}_1 - {\bf r}_2\vert}  
                 d{\bf r}_1 d{\bf r}_2 \;.
\end{align}
\end{subequations}
%
%The symbol $ \vert\vert $ is to denote the operator $ r_{12}^{-1}$ and double integration over $ {\bf r}_1 $
%and $ {\bf r}_2 $. 
Note that, while $A_{\eta\xi}$ is a typical 2\hyp{}center ERI,
$b^{(i)}_\eta$ matrix elements are not at all trivial to evaluate
because the OEP operator, which contains integration over an electron coordinate,
is present inside the double integral. Therefore, the following triple integral
has to be computed
%
\begin{equation}
 b^{(i)}_\eta = \iiint 
           \frac{ \varphi^*_\eta({\bf r}_1) \phi_i({\bf r}_2)  \rho^{\rm eff}({\bf r}_3) }
            {\vert {\bf r}_1 - {\bf r}_2 \vert \vert {\bf r}_3 - {\bf r}_2 \vert}
           d{\bf r}_1 d{\bf r}_2 d{\bf r}_3
\end{equation}
%
Computation of all the necessery integrals of this kind is very costly 
and impractical even for medium sized molecules. 
However, one can introduce the effective potential in order to eliminate one integration. 
This can be achieved by performing additional density fitting in a nearly complete intermediate basis 
%
\begin{equation}
 \hat{v}^{\rm eff} \Ket{i} = \sum_\varepsilon^{\rm Int} H_{i\varepsilon} \Ket{\varepsilon} \;,
\end{equation}
%
where the symbol `Int' denotes the intermediate basis from a chosen resolution of identity.
The working equation is therefore
%
\begin{equation}
 b^{(i)}_\eta = \sum_\varepsilon^{\rm Int} H_{i\varepsilon} R_{\varepsilon\eta} 
\end{equation}
%
with
%
\begin{equation}
 R_{\varepsilon\eta} \equiv \iint 
                       \frac{ \varphi^*_\varepsilon({\bf r}_1) \varphi_\eta({\bf r}_2) } 
                            {\vert {\bf r}_1 - {\bf r}_2\vert}  
                 d{\bf r}_1 d{\bf r}_2 \;,
\end{equation}
%
which can be easily evaluated and the matrix elements of the OEP operator found.
%Therefore, in order to use this generalized density fitting scheme
%one must to compute two\hyp{}centre electron repulsion integrals
%as well as four\hyp{}centre asymmetric electron repulsion integrals of the type $ (\alpha\beta\gamma||\eta) $.
%The evaluation of such integrals is discussed in Appendix~\ref{a:mcmurchie-davidson}.

\section{\label{s:3}Calculation Details}
\section{\label{s:4}Results and Discussion}
\subsection{\label{ss.4.1}Pauli-Repulsion Interaction Energy}
\subsection{\label{ss.4.2}Charge-Transfer Interaction Energy}
\section{\label{s:5}Summary and a few concluding remarks}

>>>>

\section{\label{s:3}Effective One-Electron Potentials: Configuration Interaction}

The time\hyp{}independent Schr{\"o}dinger equation provides the stationary solution to the
electronic state
of a molecular system in terms of its $N$\hyp{}electron wavefunction, $\Psi$,
%
\begin{equation} \label{e:schrodinger}
 E = \tBraKet{\Psi}{\mathscr{H}}{\Psi} \;,
\end{equation}
%
for a quantum Hamiltonian, $\mathscr{H} = \hat{h}(1) + \hat{o}(2)$,
with $\hat{h}(1)$ being the core Hamiltonian one\hyp{}electron operator
and $\hat{o}(2)$ the two\hyp{}electron repulsion operator.
From the above solution, any property $P$, understood as an expectation value of the
associated operator $\mathscr{P}$, can be computed according to
%
\begin{equation} \label{e:prop}
 P = \tBraKet{\Psi}{\mathscr{P}}{\Psi} \;.
\end{equation}
%
The exact solution to Eq.~\eqref{e:schrodinger} is formally given by
the configuration interaction (CI) expansion around the reference 
(approximate) wavefunction, $\Psi^{(0)}$,
typically the solution to the Hartree\hyp{}Fock (HF) problem\cite{Roothaan.RevModPhys.1951},
%
\begin{equation} \label{e:ci}
 \Ket{\Psi} = \Ket{\Psi^{(0)}} + \sum_{ra} C_{r}^{a} \Ket{\Psi_{r}^{a}} + 
	 \sum_{rsab} C_{rs}^{ab} \Ket{\Psi_{rs}^{ab}} + \ldots
\end{equation}
%
In the above equation,
$\Ket{\Psi_{rs\ldots}^{ab\ldots}}$ are called the CI configurations
with the associated CI coefficients $C_{rs\ldots}^{ab\ldots}$, created by `exciting'
one or more electrons from the $a$th ($b$th) occupied molecular orbital to the $a$th ($b$th)
virtual (unoccupied) molecular orbital, obtained from the HF solution.
By inserting Eq.~\eqref{e:ci} into Eq.~\eqref{e:prop} 
and using the Slater\hyp{}Condon rules for evaluating the one\hyp{} and two\hyp{}electron
operator matrix element between
the CI configurations, it can be shown that the following holds
%
\begin{equation} \label{e:exp-val-series}
	P =
	\sum_{ij} P_{ij} \tBraKet{i}{\hat{h}}{j}
	+ \sum_{ijkl} P_{ijkl} \BraKet{ij}{kl}  \;,
\end{equation}
%
%where $\tBraKet{i}{\hat{h}}{j} \equiv \int \phi^*_i({\bf r}) \hat{h} \phi_j({\bf r}) d{\bf r} $
%$\BraKet{ij}{kl}$ is the electron\hyp{}repulsion integral (ERI) defined by
%%
%\begin{equation}
%	\BraKet{ij}{kl} \equiv
%	\iint 
%	\frac{ \phi_i^{*}({\bf r}_1) \phi_j({\bf r}_1) 
%	       \phi_k^{*}({\bf r}_2) \phi_l({\bf r}_2) }{ \vert {\bf r}_1 - {\bf r}_2 \vert}
%	d{\bf r}_1 d{\bf r}_2  \;,
%\end{equation}
%%
where the second\hyp{}rank tensor ${\bf P}^{(2)}$ 
and the fourth\hyp{}rank tensor ${\bf P}^{(4)}$ 
are well defined and characteristic for the property of interest.
In fact, they are certain functions of the CI coefficients, i.e.,
%
\begin{equation}
{\bf P} = {\bf P}
\left(
 \left\{
  C_{r}^{a}, C_{rs}^{ab}, \ldots 
 \right\} 
\right)
\end{equation}
%
and, in principle, are known as long as the solution to the CI problem 
or its approximation is available.

@@

\subsection{\label{ss:2.3}Incorporating Electron Repulsion Integrals into Effective Potentials}

@@

\subsection{\label{ss:2.4}Definition of Interaction-Induced Property Effective Potentials}

@@

\subsection{\label{s.3344}Density Fitting of OEP's}

@@

An arbitrary one-electron potential of molecule $A$ acting on any state vector 
associated with molecule $A$ can be expanded in an auxiliary space centered 
on $A$ as
%
\begin{equation}
   \hat{v}^{A}\Ket{i} = \sum_{\xi\eta}^{\rm RI} \hat{v}^{A}\Ket{\xi} [{\bf S}^{-1}]_{\xi\eta} \BraKet{\eta}{i}
\end{equation}
%
under the necessary assumption that the auxiliary basis set is nearly complete,
i.e., 
$\sum_{\xi\eta}^{\rm RI} \Ket{\xi}[{\bf S}^{-1}]_{\xi\eta} \Bra{\eta} \cong 1$.
In the equations above, 
the `RI' symbol denotes a certain auxiliary basis set that fulfills the resolution of identity. 
%In a special case when the basis set is orthogonal (e.g., molecular orbitals)
%the above relation simplifies to
%\begin{equation}
%   v\vert i) = \sum_{\xi} v\vert \xi) ( \xi \vert i)
%\end{equation}
Such a general expansion can be obtained by 
utilizing the density fitting
in the nearly\hyp{}complete space,
%
\begin{equation}
 \hat{v}^{A}\Ket{i} = \sum_{\xi}^{\rm RI} V^A_{i\xi} \Ket{\xi} \;.
\end{equation}
%
In the above equation,
the matrix ${\bf V}^A$
is the projection of the state vector $\hat{v}^{A} \Ket{i}$
onto the complete basis $\{ \varphi_\xi \}$.
Let $Z_i[{\bf V}^A]$ be the least\hyp{}squares objective function 
%
\begin{equation} \label{e:z-compl}
 Z_i[{\bf V}^A] = \int \vert \Xi_i({\bf r}) \vert^2 d{\bf r}
\end{equation}
%
with the error density defined by
%
\begin{equation}
 \Xi_i({\bf r}) = v^A({\bf r}) \phi_i({\bf r}) - \sum_\xi^{\rm RI} V^A_{i\xi} \varphi_\xi({\bf r}) \;.
\end{equation}
%
By requiring that
%
\begin{equation} \label{e:z-necessary-requirement}
 \frac{\partial Z_i[{\bf V}^A]}{\partial V^A_{i\mu}} = 0 \text{ for all $\mu$}
\end{equation}
%
one finds the coefficients of the $i$th row of ${\bf V}^A$ to be
%
\begin{equation}
% {\bf v}_i = {\bf a}^{(i)} \cdot {\bf S}^{-1}
  V^A_{i\xi} = \sum_\eta^{\rm RI} [{\bf S}^{-1}]_{\xi\eta} a^{(i)}_\eta \;,
\end{equation}
%
where the auxiliary matrices are given by
%
\begin{subequations}
\begin{align}
 a^{(i)}_\eta &= \int \varphi^*_\eta({\bf r}) \hat{v}^A \phi_i({\bf r}) d{\bf r} \;,\\  
 S_{\eta\xi}  &= \int \varphi^*_\eta({\bf r}) \varphi_\xi({\bf r}) d{\bf r} \;.
\end{align}
\end{subequations}
%
%Since matrix elements of an OEP operator in auxiliary space can be computed 
%in the same way as the matrix elements with any other basis function, 
%one can formally write the following identity
%\begin{equation}
% (X \vert v\vert i) = \sum_{\xi\eta} S_{X \xi} [{\bf S}^{-1}]_{\xi\eta} (\eta\vert v\vert i)
%\end{equation}
%where $ X $ is an arbitrary orbital.
%When the other orbital
%does not belong to molecule $A$ but to the (changing) environment, it is 
%straightforward to compute the resulting matrix element, which is simply given as  
%\begin{equation}
%   (j_{\in B} \vert v^A \vert i_{\in A}) = \sum_\xi {S_{j\xi}} {G_{i\xi}}
%\end{equation}
%where $j$ denotes the other (environmental) basis function.
%
The working formula for $a^{(i)}_\eta$ can be found by applying 
potential form from Eq.~\eqref{e:v-eff}
along with the spectral representation of the effective density from Eq.~\eqref{e:d-spectral} 
which finally gives
%
\begin{equation}
 a^{(i)}_\eta = \sum_{x} W_{\eta i}^{(x)} + 
 \sum_{\alpha\beta} D^A_{\alpha\beta} 
  \BraKet{\alpha\beta}{\eta i} \;.
%  \iint \frac{ \phi_\alpha^*({\bf r}_1)  \phi_\beta({\bf r}_1) \varphi^*_\eta({\bf r}_2) \phi_i({\bf r}_2)}
% {\vert {\bf r}_1 - {\bf r}_2 \vert }
% d{\bf r}_1 d{\bf r}_2
\end{equation}
%

@@

Density fitting scheme from previous section has practical disadvantage of a nearly\hyp{}complete basis set
being usually very large (spanned by large amount of basis set vectors). 
Since most of basis sets used in quantum chemistry do not form a nealy complete
set, it is beneficial to design a modified scheme in which it is possible to obtain the effective 
matrix elements of the OEP operator in a incomplete auxiliary space. This can be achieved by minimizing 
the following objective function
%
\begin{equation} \label{e:z-incompl}
	Z_i[{\bf V}^A] = \iint 
        \frac{ \Xi_i^*({\bf r}_1) \Xi_i({\bf r}_2) }{\vert {\bf r}_1 - {\bf r}_2 \vert}  
         d{\bf r}_1 d{\bf r}_2  \;.
\end{equation}
%
with the error density defined by
%
\begin{equation}
 \Xi_i({\bf r}) = v^A({\bf r}) \phi_i({\bf r}) - \sum_\xi^{\rm DF} V^A_{i\xi} \varphi_\xi({\bf r}) \;.
\end{equation}
%
where the symbol `DF' denotes the generally incomplete auxiliary basis set.
From the requirement given in Eq.~\eqref{e:z-necessary-requirement}
one obtains
%
\begin{equation} \label{e:v-noncompl}
% {\bf G}^{(i)} = {\bf b}^{(i)} \cdot {\bf A}^{-1}
  V^A_{i\xi} = \sum_\eta^{\rm DF} [{\bf A}^{-1}]_{\xi\eta} b^{(i)}_\eta \;,
\end{equation}
%
where 
%
\begin{subequations}
\begin{align}
 b^{(i)}_\eta &= \iint 
                       \frac{ \varphi^*_\eta({\bf r}_1) \hat{v} \phi_i({\bf r}_2) } 
                            {\vert {\bf r}_1 - {\bf r}_2\vert}  
                 d{\bf r}_1 d{\bf r}_2 \;, \\
 A_{\eta\xi}  &= \iint 
                       \frac{ \varphi^*_\eta({\bf r}_1) \varphi_\xi({\bf r}_2) } 
                            {\vert {\bf r}_1 - {\bf r}_2\vert}  
                 d{\bf r}_1 d{\bf r}_2 \;.
\end{align}
\end{subequations}
%
%The symbol $ \vert\vert $ is to denote the operator $ r_{12}^{-1}$ and double integration over $ {\bf r}_1 $
%and $ {\bf r}_2 $. 
Note that, while $A_{\eta\xi}$ is a typical 2\hyp{}center ERI 
that can be evaluated by standard means,
$b^{(i)}_\eta$ matrix elements are not at all trivial to calculate
because the OEP operator, which contains integration over an electron coordinate,
is present inside the double integral. Therefore, the following triple integral,
%
\begin{equation} \label{e:triple-integral}
 b^{(i)}_\eta = \iiint 
           \frac{ \varphi^*_\eta({\bf r}_1) \phi_i({\bf r}_2)  \rho^{\rm eff}({\bf r}_3) }
            {\vert {\bf r}_1 - {\bf r}_2 \vert \vert {\bf r}_3 - {\bf r}_2 \vert}
           d{\bf r}_1 d{\bf r}_2 d{\bf r}_3 \;,
\end{equation}
%
has to be computed.
Obtaining all the necessery integrals of this kind directly 
by performing integrations in Eq.~\eqref{e:triple-integral} is very costly 
and impractical even for medium sized molecules. 
However, one can introduce the effective potential in order to eliminate one integration. 
This can be achieved by performing additional density fitting in a nearly complete intermediate basis 
%
\begin{equation}
 \hat{v}^{\rm eff} \Ket{i} = \sum_\varepsilon^{\rm RI} H_{i\varepsilon} \Ket{\varepsilon} \;,
\end{equation}
%
%where the symbol `RI' denotes the intermediate basis from a chosen resolution of identity.
The working equation is therefore given by
%
\begin{equation}
 b^{(i)}_\eta = \sum_\varepsilon^{\rm Int} H_{i\varepsilon} R_{\varepsilon\eta} 
\end{equation}
%
with
%
\begin{equation}
 R_{\varepsilon\eta} \equiv \iint 
                       \frac{ \varphi^*_\varepsilon({\bf r}_1) \varphi_\eta({\bf r}_2) } 
                            {\vert {\bf r}_1 - {\bf r}_2\vert}  
                 d{\bf r}_1 d{\bf r}_2 \;,
\end{equation}
%
which can be easily evaluated and the matrix elements of the OEP operator found
from Eq.~\eqref{e:v-noncompl}.
%Therefore, in order to use this generalized density fitting scheme
%one must to compute two\hyp{}centre electron repulsion integrals
%as well as four\hyp{}centre asymmetric electron repulsion integrals of the type $ (\alpha\beta\gamma||\eta) $.
%The evaluation of such integrals is discussed in Appendix~\ref{a:mcmurchie-davidson}.

It is emphasized here that, as long as the state vector $\Ket{i}$, OEP operator $\hat{v}^A$ 
and the auxiliary and intermediate basis sets depend solely on one unperturbed molecule $A$, the matrix elements
$V^A_{i\xi}$ can be calculated just once and stored in a file as effective fragment parameters.

\section{\label{s:4}Calculation Details}

All the models that were used to test the theory presented in this work
were implemented in our in\hyp{}house plugin to {\sc Psi4} quantum chemistry program.\cite{Psi4.JCTC.2017}

\section{\label{s:5}Results and Discussion}
\subsection{\label{ss.5.1}Pauli-Repulsion Interaction Energy}
\subsection{\label{ss.5.2}Charge-Transfer Interaction Energy}

Charge transfer (CT) between molecules occurs when the net electronic populations 
of interacting molecules change which leads to an additional stabilization 
of a molecular aggregate. Since the effect of CT on the interaction energy 
is known to be often non\hyp{}negligible, it is quite important to model 
CT energies with reasonable accuracy and efficiency when performing energy 
and molecular dynamics calculations with \emph{ab initio} force fields, 
especially in donor\hyp{}acceptor systems such as H\hyp{}bonded species. 
However, estimating the CT stabilization energy is usually much more expensive 
than estimating the contributions due to other effects such as electrostatics, 
exchange\hyp{}repulsion and dispersion.\cite{Gordon.Fedorov.Pruitt.Slipchenko.ChemRev.2012}

As an illustration of application of the OEP technique for efficient calculation 
of CT energy we consider the second\hyp{}order perturbation theory of Murrell et al.,
in which the energy associated with the charge transfer from molecule $A$ to $B$ is given by
%
\begin{equation} \label{e:ct-murell-etal.e}
 E^{A\rightarrow B} = 2 \sum_{i\in A}^{\rm Occ} \sum_{n\in B}^{\rm Vir} 
  \frac{\vert V^{A\rightarrow B}_{in} \vert^2 }{\varepsilon_i - \varepsilon_n}
\end{equation}
%
%coupling constants between occupied and virtual molecular orbitals 
%are re-expressed by the effective one-electron potentials. The resulting formula provides new possibilities 
%of optimization and applications in condensed phase simulations. [3]
%
where $\varepsilon_i$ and $\varepsilon_n$ are the canonical energies
of the occupied (denoted by `Occ') and virtual (denoted by `Vir') 
HF molecular orbitals, respectively,
whereas the coupling constant is given by
%
\begin{multline} \label{e:ct-murell-etal.vin}
 V^{A\rightarrow B}_{in} = 
        \tBraKet{i}{\hat{v}^B_{\rm ch} }{n} 
      - \sum_{j\in B}^{\rm Occ} \BraKet{nj}{ij} 
      - \sum_{k\in A}^{\rm Occ} S_{nk} \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} \\
      - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{i} }{n}  
     + \sum_{k\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}  
        S_{kj} \left( 1 - \delta_{ik} \right) 
        \BraKet{nj}{ik} \;.
\end{multline}
%
In the above expression, the following effective potentials,
%
\begin{subequations} 
\begin{align} \label{e:ct-murell-etal.vtot-vi}
 \hat{v}^B_{\rm tot} &= \hat{v}^B_{\rm nuc} + 2\sum_{j\in B}^{\rm Occ} \hat{v}^B_{jj} \quad\text{ and }\\ 
 \hat{v}^A_{i      } &= \hat{v}^A_{\rm tot} - 2\hat{v}^A_{ii} \;,
\end{align}
\end{subequations}
%
were introduced without making any approximation to the original equation
from Ref. One can immediately notice that the five summation terms
from Eq.~\eqref{e:ct-murell-etal.vin} can be classified into three groups
regarding the type of ERI's that are required:
(i) $\BraKet{AB}{BB}$ -- the first two terms;
(ii) $\BraKet{AA}{BB}$ -- the third term and
(iii) $\BraKet{BB}{AA}$ -- the two last terms. 
%Note that groups (ii) and (iii) correspond to the Coulomb interactions between
%The group (i) corresponds to the overlap\hyp{}like integrals, whereas the groups (ii) and (iii)
%correspond to the Coulomb\hyp{}like integrals

\paragraph{Group (i).}
Group (i) can be rewritten by noticing that
%
\begin{equation} \label{e:ct-murell-etal.group-i.notice}
 \sum_{j\in B}^{\rm Occ} 
 \BraKet{nj}{ij} =
 \Bra{i} \left[ \hat{v}^B_{nj} \Ket{j} \right]  \;,
\end{equation}
%
which, by combining with the first term from Eq.~\eqref{e:ct-murell-etal.vin}, 
allows to apply the density fitting in an
incomplete basis set as follows
%
\begin{equation}
\Bra{i} \left[ \hat{v}^B_{\rm tot} \Ket{n} - \sum_{j\in B}^{\rm Occ} \hat{v}^B_{nj} \Ket{j} \right]
\cong \Bra{i} \sum_{\eta\in B}^{\rm DF} V^B_{n\eta} \Ket{\eta}
\end{equation}
%
with the error density defined by
%
\begin{multline}
 \Xi_n({\bf r}) = 
  \left\{
      v_{nuc}^B({\bf r}) + 2 \sum_{j\in B}^{\rm Occ}
   \int \frac{ \vert \phi_j({\bf r}') \vert^2 }{\vert {\bf r}' - {\bf r} \vert} d{\bf r}'
  \right\}
  \phi_n({{\bf r}})  \\
 - \sum_{j\in B}^{\rm Occ} \phi_j({{\bf r}})
   \int \frac{ \phi^*_n({\bf r}') \phi_j({{\bf r}'}) }{\vert {\bf r}' - {\bf r} \vert} d{\bf r}'
 - \sum_{\eta\in B}^{\rm DF} V^B_{n\eta} \varphi_\eta({\bf r}) \;.
\end{multline}
%
Substituting the above equation into Eq.~\eqref{e:z-incompl}
%and then
%using resolution of identity to simplify the 
leads to
%
\begin{equation}
 V^B_{n\xi} = \sum_{\eta\in B}^{\rm DF} \left[ {\bf A}^{-1} \right]_{\xi\eta} b_\eta^{(n)} \;,
\end{equation}
%
where
%
\begin{multline}
 b^{(n)}_\eta = \iint 
           d{\bf r}_1 d{\bf r}_2  
           \frac{ \varphi^*_\eta({\bf r}_1) }
            {\vert {\bf r}_1 - {\bf r}_2 \vert } \\ \times
          \left[ 
           \hat{v}^B_{\rm tot}  \phi_n({\bf r}_2) 
         - \sum_{j\in B}^{\rm Occ} \hat{v}^B_{nj} \phi_j({{\bf r}_2})
           \right]  \;.
\end{multline}
%
The above result is given by a sum of triple integrals from Eq.~\eqref{e:triple-integral}.
However, application of the resolution of identity as follows
%
\begin{equation}
 \hat{v}^B_{\rm tot} \Ket{n} - \sum_{j\in B}^{\rm Occ} \hat{v}^B_{nj} \Ket{j}
 \cong \sum_{\varepsilon\in B}^{\rm RI} H_{n\varepsilon}^B \Ket{\varepsilon}
\end{equation}
%
leads to a much simpler formula
that requires only the one- and two\hyp{}electron integrals. 
Therefore, by combining
equations %which equations?
one arrives to
%
\begin{equation}
 V^B_{n\xi} = \sum_{\eta\in B}^{\rm DF} 
          \sum_{\varepsilon\in B}^{\rm RI}
         \left[ {\bf A}^{-1} \right]_{\xi\eta}
         H_{n\varepsilon}^B R_{\varepsilon \eta} \;,
\end{equation}
%
where
%
\begin{equation}
 H_{n\varepsilon}^B = \sum_{\eta\in B}^{\rm RI} \left[ {\bf S}^{-1} \right]_{\varepsilon\eta}
   a_\eta^{(n)}
\end{equation}
%
and
%%
\begin{align}
 a^{(n)}_\eta &= \tBraKet{\eta}{\hat{v}^B_{\rm tot}}{n}
      - \sum_{j\in B}^{\rm Occ} \tBraKet{\eta}{\hat{v}^B_{nj}}{j} \nonumber \\
 &= \sum_{y\in B} W^{(y)}_{\eta n} 
  + \sum_{j\in B}^{\rm Occ} 
  \left\{
   2\BraKet{\eta n}{jj} - \BraKet{\eta j}{nj} 
  \right\} \;.
\end{align}
%
Note that all the calculations that are required to obtain $V^B_{n\xi}$ are performed
solely on the densities and basis sets associated with the unperturbed molecule $B$.
Therefore, $V^B_{n\xi}$ can be considered as effective fragment parameters
used to compute the first two terms of Eq.~\eqref{e:ct-murell-etal.vin} by
%
\begin{equation}
        \tBraKet{i}{\hat{v}^B_{\rm ch} }{n} 
      - \sum_{j\in B}^{\rm Occ} \BraKet{nj}{ij} 
       = \sum_{\eta\in B}^{\rm RI} V^B_{n\eta} S_{\eta i} \;,
\end{equation}
%
which is a great simplification over the original form of group (i)
because only the overlap integrals between the $i$th MO on molecule $A$
and $\eta$th auxiliary orbital on molecule $B$ need to be evaluated on the fly.

\paragraph{Group (ii).}
\paragraph{Group (iii).}
\paragraph{Final OEP-based form of the coupling constant.}
Gathering the results from Eqs. the coupling constant
can be rewritten as follows
%
\begin{equation} \label{e:ct-murell-etal.vin.oep}
 V_{in}^{A\rightarrow B}
       = \sum_{\eta\in B}^{\rm RI} V^B_{n\eta} S_{\eta i} \;.
\end{equation}
%
Note that the only approximations made on it so fare was the application of density fitting
and resolution of identity. If the associated basis sets are sufficiently large, the errors
due to this approximation can be minimal and negligible.
